---
title: "Capital One Project"
author: "Jerry Liu"
date: "October 2018"
output: github_document
---

### Load libraries

```{r load}
library("tidyverse")
```


### Loading data
First, we load the data in the bike data from the csv file and also rename some columns for easier use in later analysis.

```{r load-data}

bikes <- read_csv("metro-bike-share-trip-data.csv")
colnames(bikes)[colnames(bikes)=="Starting Station Latitude"] <- "slat"
colnames(bikes)[colnames(bikes)=="Starting Station Longitude"] <- "slong"
colnames(bikes)[colnames(bikes)=="Starting Station ID"] <- "ssid"
colnames(bikes)[colnames(bikes)=="Ending Station ID"] <- "esid"
colnames(bikes)[colnames(bikes)=="Duration"] <- "duration"
colnames(bikes)[colnames(bikes)=="Passholder Type"] <- "passholdertype"

```

### Loading data
First, we load the data in the bike data from the csv file and also rename some columns for easier use in later analysis.

```{r duration}
x = bikes %>%
  group_by(passholdertype) %>%
  summarise_at(vars(duration), funs(mean(., na.rm=TRUE)))
x$duration = x$duration/60

ggplot(x, aes(passholdertype, duration)) + geom_col() + labs(title="The average duration of use between different Pass Holders") + labs(x = "Pass Holder Type") + labs(y = "Average Duration (Minutes)")

```

### What are the most popular start/stop stations?

Let's take a look at the most popular start/stop stations. In order to do this, I first grouped by start and end station id's and then counted the number of occurences of those id's.
```{r startstop}
MostFreqentStart = count(bikes, ssid) %>%
  arrange(desc(n)) %>%
  head(5) 
names(MostFreqentStart) <- c("Starting Station ID", "Frequency")
MostFreqentEnd = count(bikes, esid) %>%
  arrange(desc(n)) %>%
  head(5) 
names(MostFreqentEnd) <- c("Ending Station ID", "Frequency")
MostFreqentEnd
names(MostFreqentStart) <- c("Starting Station ID", "Frequency")
MostFreqentStart

```

### What is the average distance traveled?
```{r avgdistance}
bikes = bikes %>%
  filter(duration>1)
mean(bikes$duration)
1555.302
```
Here, we can see that the mean amount of time people spent on their bikes was 1555 seconds, which is about 26 minutes. If we assume these people were biking at a average pace of around 10mph, we can conclude that the average trip length was around 26/60*10 miles, or `4.3` miles.

### How many riders include bike sharing as a regular part of their commute?
```{r regcommute}
RegCommute = count(bikes, passholdertype) %>%
  arrange(desc(n)) %>%
  head(5)
names(RegCommute) <- c("Pass Type", "# of people using")
RegCommute
```
Let's assume that people who use the walk-up pass aren't regular users and therefore don't use it as part of their commute. Summing up the people who bought passes, 81304+41224+382 =  `122,910`. Thus, we conclude that `(122,910)/(122,910+41224)` is the % of people using the bikes for their regular commute. This evalulates to around 74.8% of users. 
