---
title: "Capital One Project"
author: "Jerry Liu"
date: "October 2018"
output: github_document
---

### Load libraries

For this project, we'll be using the TidyVerse package to help with data analysis.
```{r load, echo=FALSE, results='hide',message=FALSE}
options(warn=-1)
library("tidyverse")
```


### Loading data

First, we load the data in the bike data from the csv file and also rename some columns for easier use in later analysis.

```{r load-data, echo=FALSE, results='hide',message=FALSE}

bikes <- read_csv("metro-bike-share-trip-data.csv")
colnames(bikes)[colnames(bikes)=="Starting Station Latitude"] <- "slat"
colnames(bikes)[colnames(bikes)=="Starting Station Longitude"] <- "slong"
colnames(bikes)[colnames(bikes)=="Starting Station ID"] <- "ssid"
colnames(bikes)[colnames(bikes)=="Ending Station ID"] <- "esid"
colnames(bikes)[colnames(bikes)=="Duration"] <- "duration"
colnames(bikes)[colnames(bikes)=="Start Time"] <- "starttime"
colnames(bikes)[colnames(bikes)=="Passholder Type"] <- "passholdertype"

```


### Investigating how Bike Usage Changes over the Seasons

Here we graph the number of bike uses each day. We can see that bike use is cyclical, with usage peaking in the summer and at its low points during the winter.

```{r days, echo=FALSE, results='hide',message=FALSE}

z = bikes
z$day <- format(as.POSIXct(z$starttime), "%Y-%m-%d")
z$day = as.POSIXct(z$day)
z = z %>%
  group_by(day) %>%
  count(day)
plot = ggplot(z, aes(day, n)) 
plot = plot + 
  labs(title="Distribution of Passholders using Bikes") + labs(x = "Day") + labs(y = "Count (# of uses)") + geom_point( color="blue")
plot

```

### Types of Bike Users

Here, we can see that most people using the bikes are monthly pass users, but there are still a decent amount of people using walk up.

```{r passbikes}

y = count(bikes, passholdertype)
y
ggplot(y, aes(passholdertype, n)) + geom_col() 

```

### Investigating how usage of bikes changes

This graph shows something very interesting, in that people using the bikes as walk-ups use them on average 2-3x longer than monthly or flex pass holders.

```{r duration}
x = bikes %>%
  group_by(passholdertype) %>%
  summarise_at(vars(duration), funs(mean(., na.rm=TRUE)))
x$duration = x$duration/60

ggplot(x, aes(passholdertype, duration)) + geom_col() + labs(title="The average duration of use between different Pass Holders") + labs(x = "Pass Holder Type") + labs(y = "Average Duration (Minutes)")

```

### What are the most popular start/stop stations?

Let's take a look at the most popular start/stop stations. In order to do this, I first grouped by start and end station id's and then counted the number of occurences of those id's.

```{r startstop}
MostFreqentStart = count(bikes, ssid) %>%
  arrange(desc(n)) %>%
  head(5) 
names(MostFreqentStart) <- c("Starting Station ID", "Frequency")
MostFreqentEnd = count(bikes, esid) %>%
  arrange(desc(n)) %>%
  head(5) 
names(MostFreqentEnd) <- c("Ending Station ID", "Frequency")
MostFreqentEnd
names(MostFreqentStart) <- c("Starting Station ID", "Frequency")
MostFreqentStart

```

### What is the average distance traveled?

```{r avgdistance}
bikes = bikes %>%
  filter(duration>1)
mean(bikes$duration)
1555.302
```
Here, we can see that the mean amount of time people spent on their bikes was 1555 seconds, which is about 26 minutes. If we assume these people were biking at a average pace of around 10mph, we can conclude that the average trip length was around 26/60*10 miles, or `4.3` miles.

### How many riders include bike sharing as a regular part of their commute?

```{r regcommute}
RegCommute = count(bikes, passholdertype) %>%
  arrange(desc(n)) %>%
  head(5)
names(RegCommute) <- c("Pass Type", "# of people using")
RegCommute
```
Let's assume that people who use the walk-up pass aren't regular users and therefore don't use it as part of their commute. Summing up the people who bought passes, 81304+41224+382 =  `122,910`. Thus, we conclude that `(122,910)/(122,910+41224)` is the % of people using the bikes for their regular commute. This evalulates to around 74.8% of users. 
